### 2025-01-27 week3/main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Welcome to Overleaf --- just edit your LaTeX on the left,
% and we'll compile it for you on the right. If you open the
% 'Share' menu, you can invite other users to edit at the same
% time. See www.overleaf.com/learn for more info. Enjoy!
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{article}
\usepackage[
backend=biber,
style=ieee,
]{biblatex}
\title{Annotated Bibliography}
\addbibresource{sample.bib} %Imports bibliography file

\begin{document}

\maketitle

All citations are in IEEE Style, as I want some foray into writing academic papers in
the computer science field.

% A brief annotation (150-200 words) which does all four of the following:
% Identifies the purpose/s of the source (What change do you think the author hoped would happen when people read this source?) 
% Explains who you think the intended audience of the source is, and why you think this
% Explains some key strengths of the source and how it can be useful as you investigate your topic
% Explains possible limitations or weaknesses you see in the source
% (eg. credibility, scope of information, author bias, etc.) Keep in mind
% that you might have to do a little additional research to get some other
% perspectives on your source to evaluate its reliability (eg. book reviews,
% author bios, original publication dates etc.)

\section{Book}

Citation: \cite{stokes2021}
Link: https://ucsc.primo.exlibrisgroup.com/permalink/01CDL_SCR_INST/15r5l0d/alma991025892758804876

\section{Newspaper}

Citation: \cite{thomas2024}
Link: https://ucsc.primo.exlibrisgroup.com/permalink/01CDL_SCR_INST/ojisf2/cdi_proquest_newspapers_2937453324

\section{Paper}

Citation: \cite{ma2024}
Link: https://ucsc.primo.exlibrisgroup.com/permalink/01CDL_SCR_INST/ojisf2/cdi_ieee_primary_10440501

\section{Video}

Citation: \cite{bloomberglaw2023}
Link: https://www.youtube.com/watch?v=bRqwTP2eKJY

\medskip

\printbibliography
\end{document}